---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
set.seed(2022)
```

# ggautomap

<!-- badges: start -->
[![ggautomap status badge](https://cidm-ph.r-universe.dev/badges/ggautomap)](https://cidm-ph.r-universe.dev)
<!-- badges: end -->

`ggautomap` provides `{ggplot2}` geometries that make use of `{cartographer}`, a
framework for matching place names with map data. With `ggautomap` your input
dataset doesn't need to be spatially aware: the geometries will automatically
attach the map data (providing it's been registered with `{cartographer}`).

## Installation

You can install ggautomap like so:

``` r
options(repos = c(
  cidmph = 'https://cidm-ph.r-universe.dev',
  CRAN = 'https://cloud.r-project.org'))
  
install.packages('ggautomap')
```

## Example

Let's use the example dataset from `{cartographer}`:

```{r}
library(cartographer)

head(nc_type_example)
```

A possible workflow is to use [`cartographer::add_geometry()`] to convert this
into a spatial data frame and then use [`ggplot2::geom_sf()`] to draw it.

`ggautomap` instead provides geoms that do this transparently as needed, so you
don't need to do a lot of boilerplate to wrangle the data into the right form
before handing it off to the plotting code.

You can then use this data with the geoms to provide map summaries. For the map
outline, you'll need to specify `feature_type`, but the other geoms will usually
be able to guess the correct value.

```{r example-basic, fig.width=8, fig.height=3}
library(ggplot2)
library(ggautomap)

nc_type_example2 <- cartographer::nc_type_example[1:49,]

ggplot(nc_type_example2, aes(location = county)) +
  geom_boundaries(feature_type = "sf.nc") +
  geom_geoscatter(aes(colour = type), size = 0.5)
ggplot(nc_type_example2, aes(location = county)) +
  geom_boundaries(feature_type = "sf.nc") +
  geom_centroids(aes(colour = type, scale = 6), size = 0.5)
ggplot(nc_type_example2, aes(location = county)) +
  geom_boundaries(feature_type = "sf.nc") +
  geom_choropleth() +
  scale_fill_steps(low = "#e6f9ff", high = "#00394d", na.value = "white")
ggplot(nc_type_example2, aes(location = county)) +
  geom_boundaries(feature_type = "sf.nc") +
  geom_pie(aes(fill = type), pie_radius = 0.1)
```

