---
title: "Getting started with 'ggautomap'"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with 'ggautomap'}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(nswgeo)
library(ggautomap)
library(ggplot2)
library(dplyr, warn.conflicts = FALSE)
```

This article provides some recipes for plots that might be of interest. These
examples use map data from the `{nswgeo}` package.

Many of the examples use the same example dataset modelled after the structure
of a linelist: rows are distinct events, and they can have a `type` of `A` or `B`.
Each event is associated with a location described at different granularities by
the `postcode`, `lga`, and `lhd` columns.

```{r}
head(covid_cases_nsw)
```

You need to specify which column has the feature by setting the `location`
aesthetic. This example has three different columns of locations for different
feature types; your dataset only needs to have one of these.

In general you'll start with `geom_boundaries()` to draw the base map. This geom
needs to be told which `feature_type` you're after (e.g. `"nswgeo.lga"` for LGAs).
All of the summary geoms of `ggautomap` can then be used to draw your data.

## Scatter

```{r scatter, fig.height = 3, fig.width = 7}
covid_cases_nsw %>%
  ggplot(aes(location = lga)) +
  geom_boundaries(feature_type = "nswgeo.lga") +
  geom_geoscatter(aes(colour = type), sample_type = "random", size = 0.5) +
  coord_automap(feature_type = "nswgeo.lga", xlim = c(147, 153), ylim = c(-33.7, -29)) +
  guides(colour = guide_legend(override.aes = list(size = 1))) +
  theme_void()
```

Points are drawn at random within the boundaries of their location.


## Insets

To show a zoomed in part of the map as an inset, you can configure an inset and
provide it to each relevant geom. The geoms in this package are all inset-aware.
See `{ggmapinset}` for details.

```{r inset, fig.height = 3, fig.width = 7}
covid_cases_nsw %>%
  ggplot(aes(location = lga)) +
  geom_boundaries(feature_type = "nswgeo.lga") +
  geom_geoscatter(aes(colour = type), size = 0.5) +
  geom_inset_frame() +
  coord_automap(feature_type = "nswgeo.lga", inset = configure_inset(
    centre = "Blacktown", radius = 40, units = "km",
    scale = 7, translation = c(400, -100)
  )) +
  theme_void()
```

## Packed points

This next example uses `geom_centroids()` to place the points in a
packed circle in the centre of each feature. It also shows how you can
fine-tune the plot with the usual `{ggplot2}` functions.

```{r packed, fig.height = 3, fig.width = 7}
covid_cases_nsw %>%
  dplyr::filter(year >= 2021) %>%
  ggplot(aes(location = lhd)) +
  geom_boundaries(feature_type = "nswgeo.lhd") +
  geom_centroids(aes(colour = type), position = position_circle_repel_sf(scale = 35), size = 1) +
  geom_inset_frame() +
  coord_automap(feature_type = "nswgeo.lhd", inset = configure_inset(
    centre = "Sydney", radius = 80, units = "km", feature_type = "nswgeo.lhd",
    scale = 6, translation = c(650, -100)
  )) +
  facet_wrap(vars(year)) +
  labs(x = NULL, y = NULL) +
  theme_void() +
  theme(strip.text = element_text(size = 12))
```


## Choropleths

If your data has multiple rows for each location (such as our example dataset where
the rows are disease cases) then you can use `geom_choropleth()` to aggregate these
into counts.

```{r choro-long, fig.height = 3, fig.width = 7}
covid_cases_nsw %>%
  ggplot(aes(location = lhd)) +
  geom_choropleth() +
  geom_boundaries(
    feature_type = "nswgeo.lhd", colour = "black", linewidth = 0.1,
    outline.aes = list(colour = NA)
  ) +
  geom_inset_frame() +
  coord_automap(feature_type = "nswgeo.lhd", inset = configure_inset(
    centre = "Western Sydney", radius = 60, units = "km",
    scale = 5, translation = c(400, -100)
  )) +
  scale_fill_steps(low = "#e6f9ff", high = "#00394d", n.breaks = 5, na.value = "white") +
  theme_void()
```

On the other hand, if your dataset has only one row per location and there is an
existing column that you'd like to map to the `fill` aesthetic, then instead use
`geom_sf_inset(..., stat = "automap")`:

```{r choro-wide, fig.height = 3, fig.width = 7}
summarised_data <- data.frame(
  lhd = c("Western Sydney", "Sydney", "Far West", "Mid North Coast", "South Western Sydney"),
  cases = c(250, 80, 20, NA, 100)
)

summarised_data %>%
  ggplot(aes(location = lhd)) +
  geom_sf_inset(aes(fill = cases), stat = "automap", colour = NA) +
  geom_boundaries(
    feature_type = "nswgeo.lhd", colour = "black", linewidth = 0.1,
    outline.aes = list(colour = NA)
  ) +
  geom_inset_frame() +
  coord_automap(feature_type = "nswgeo.lhd", inset = configure_inset(
    centre = "Western Sydney", radius = 60, units = "km",
    scale = 3.5, translation = c(350, 0)
  )) +
  scale_fill_gradient(low = "#e6f9ff", high = "#00394d", na.value = "grey90") +
  theme_void()
```


## Positioning text

These examples give some different ways of placing text, accounting for possible insets.

```{r text-inset, fig.height = 3, fig.width = 7}
covid_cases_nsw %>%
  ggplot(aes(location = lhd)) +
  geom_choropleth() +
  geom_boundaries(feature_type = "nswgeo.lhd") +
  geom_inset_frame() +
  geom_sf_label_inset(aes(label = lhd),
    stat = "automap_coords",
    data = ~ dplyr::slice_head(.x, by = lhd)
  ) +
  coord_automap(feature_type = "nswgeo.lhd", inset = configure_inset(
    centre = "Western Sydney", radius = 60, units = "km",
    scale = 3.5, translation = c(350, 0)
  )) +
  labs(x = NULL, y = NULL) +
  theme_void()
```

The repulsive labels from `{ggrepel}` can be used; they just require a bit of
massaging since they don't natively understand the spatial data. Note that
you may also wish to use `point.padding = NA` to disable the default repulsion
caused by the labelled points, which is good for labelling scatter plots but
often doesn't make sense in mapping contexts.

```{r text-repel, fig.height = 4, fig.width = 7}
library(ggrepel)

# label all features that have data
covid_cases_nsw %>%
  ggplot(aes(location = lhd)) +
  geom_choropleth() +
  geom_boundaries(feature_type = "nswgeo.lhd") +
  geom_inset_frame() +
  geom_label_repel(
    aes(
      x = after_stat(x_inset),
      y = after_stat(y_inset),
      label = lhd
    ),
    stat = "automap_coords",
    nudge_x = 3,
    nudge_y = 1,
    point.padding = NA,
    data = ~ dplyr::slice_head(.x, by = lhd)
  ) +
  scale_fill_distiller(direction = 1) +
  coord_automap(feature_type = "nswgeo.lhd", inset = configure_inset(
    centre = "Western Sydney", radius = 60, units = "km",
    scale = 3.5, translation = c(350, 0)
  )) +
  labs(x = NULL, y = NULL) +
  theme_void()

# label all features in the map regardless of data, hiding visually overlapping labels
covid_cases_nsw %>%
  ggplot(aes(location = lhd)) +
  geom_choropleth() +
  geom_boundaries(feature_type = "nswgeo.lhd") +
  geom_inset_frame() +
  geom_label_repel(
    aes(
      x = after_stat(x_inset),
      y = after_stat(y_inset),
      geometry = geometry,
      label = lhd_name
    ),
    stat = "sf_coordinates_inset",
    data = cartographer::map_sf("nswgeo.lhd"),
    point.padding = NA,
    inherit.aes = FALSE
  ) +
  scale_fill_distiller(direction = 1, palette = 2) +
  coord_automap(feature_type = "nswgeo.lhd", inset = configure_inset(
    centre = "Western Sydney", radius = 60, units = "km",
    scale = 4, translation = c(500, 0)
  )) +
  labs(x = NULL, y = NULL) +
  theme_void()
```

## Loading custom maps and shape files

ggautomap accesses map data via the cartographer package's registry. See `vignette("registering_maps", package = "cartographer")` for a guide on how to register new map data. This can be as simple as a one-liner calling `cartographer::register_map()` to assign map data to a name that ggautomap can use.

Before proceeding to load custom shape files, check the R package [rnaturalearth](https://github.com/ropensci/rnaturalearth) which contains maps for world maps and countries.

Map data needs to be in any format understood by `sf::read_sf()`, such as a shape file (`.shp`).
For example Australian maps can be retrieved from [ABS Digital boundary files](https://www.abs.gov.au/statistics/standards/australian-statistical-geography-standard-asgs-edition-3/jul2021-jun2026/access-and-downloads/digital-boundary-files).

First read in the shape file using the `sf` package:

```{r, eval=FALSE}
sf <- sf::read_sf("map_data.shp")
```
Shape files can be subset in a similar fashion to data.frames using `dplyr`:

```{r, eval=FALSE}
# subset to only include Victorian postcodes only
sf_vic <-
  sf::read_sf(map_files) |>
  mutate(postcode = as.integer(POA_CODE21)) |>
  filter(POA_CODE21 >= 3000 & POA_CODE21 <= 3999)
```

After reading in the shape file, register with `cartographer` package:
```{r, eval=FALSE}
cartographer::register_map(
  "sf.vic", # the map name
  data = sf_vic, # this is the object we subsetted above
  feature_column = "postcode", # data column to include
)
```

Check the map has been registered:
```r
cartographer::feature_types()
 [1] "maps.italy"                    "rnaturalearth.countries_hires" "sf.vic"                        "maps.lakes"
 [5] "rnaturalearth.countries"       "rnaturalearth.australia"       "maps.nz"                       "maps.world"
 [9] "maps.state"                    "sf.nc"                         "maps.france"
```

To use the registered shapefile:
```{r, eval=FALSE}
ggplot() +
  geom_boundaries(feature_type = "sf.vic") +
  theme_void()
```
